---
title: "mergeNDandInequity"
output: html_document
date: "2025-07-08"
---

The mergeNDandInequity.RMD code was used to map the time-series ND GAIN data to each of the 800,000+ points filtered in the contextual inequity dataset. While doing this would still cause every point in a country to have the same ND GAIN score, this allows for a point-level resolution for this dataset as well. This markdown file was also used to create a dataset with regional averages for all scores.

```{r setup, include=FALSE}

library(shiny)
library(leaflet)
library(terra)
library(bslib)
library(shiny)
library(leaflet)
library(dplyr)
library(sf)
library(viridis)
library(tidyverse)
library(qs)
library(pryr)
library(rnaturalearth)
library(shinyjs)
library(ggthemes)
library(plotly)

gain <- readRDS("/Users/student/Desktop/Data+/climate_plus_summer_2025/June16Dashboard/data/gain_coastal_filtered.rds")
ineq <- readRDS("/Users/student/Desktop/Data+/climate_plus_summer_2025/June16Dashboard/data/inequity_filtered5k.rds")
wide <- readRDS("/Users/student/Desktop/Data+/climate_plus_summer_2025/June16Dashboard/data/wide_global_temp_anomaly_1990_2025 copy 2.rds")
```

Creating a dataset where data points were grouped by region (NAME_2), and indicator scores were averaged across the points in each regino.

```{r}

# Create inequity dataset without geometry.

ineq_no_geo <- ineq %>%
  st_drop_geometry()

# Create regional dataset with regional averages.

reg <- ineq %>%
  group_by(NAME_2) %>%
  summarize(
    ID = first(ID),
    name_en = first(name_en),
    iso_a3.x = first(iso_a3.x),
    NAME_1 = first(NAME_1),
    across(mean.count.grav.V2.log.sc:hierachical.score.rank.ineq, mean, na.rm = TRUE),
    .groups = "drop"
  )

```

SaveRDS file.

```{r}
filepath = "/Users/student/Desktop/Data+/climate_plus_summer_2025/June16Dashboard/data/regional_scores_of_inequity_filtered_5k.rds"
reg %>%
  st_drop_geometry()

saveRDS(reg, filepath)

```

Create new wide version of gain to be combined with contextual inequity dataset 800,000+ points. column names were organized by score and year.

```{r}

gain_wide <- gain %>%
  pivot_wider(
    id_cols = c(ISO3, Name),
    names_from = Year,
    values_from = -c(ISO3, Name, Year),
    names_sep = "__"
  )

filepath = "/Users/student/Desktop/Data+/climate_plus_summer_2025/June16Dashboard/data/gain_wide.rds"

saveRDS(gain_wide, filepath)

```

Assign GAIN scores and indicators to each of the 800,000+ points according to country.

```{r}

joined <- full_join(ineq, gain_wide, by = join_by(iso_a3.x == ISO3))
joined <- joined %>%
  select(-c(4:40))

# Removed inequity indicators to make file smaller (and contextual inequity dataset has its own file anyways).

```

Save it to interactive dashboard's data folder.

```{r}

filepath = "/Users/student/Desktop/Data+/climate_plus_summer_2025/June16Dashboard/data/gain_wide_points.rds"

saveRDS(joined, filepath)

```

