---
title: "Global Temperature Anomaly Filtered"
output: html_document
date: "2025-06-17"
---

```{r setup, include=FALSE}
library(tidyverse)
library(tidyr)
library(dplyr)
library(sf)
library(ggplot2)
library(rnaturalearth)
library(rnaturalearthdata)
library(lubridate)

all_times <- read.csv("/Users/student/Downloads/Global_Monthly_Temperature_Anomaly_(1850_-_present).csv")

all_times <- all_times %>%
  mutate(
    # convert to proper datetime format
    datetime_obj = ymd_hms(Date_),
    
    # extract components
    year = year(datetime_obj),
    month = month(datetime_obj),
    day = day(datetime_obj),
    hour = hour(datetime_obj),
    minute = minute(datetime_obj),
    second = second(datetime_obj),
    timezone = tz(datetime_obj)
  ) %>%
  
  # remove the temporary datetime object if needed
  select(-datetime_obj)


```

Remove unnecessary variables.

```{r}

all_times <- all_times %>%
  select(-Date_) %>%
  select(-minute) %>%
  select(-second)

```

Filter to only between 1990 and 2025

```{r}

years <- 1990:2025

all_times <- all_times %>%
  filter(year %in% years)

```



Convert to sf object.

```{r}

points_sf <- all_times %>%
  st_as_sf(coords = c("X", "Y"), crs = 4326)

```


Initialize coastline variables.

```{r}


coastline <- ne_download(scale = "small", type = "coastline", category = "physical", returnclass = "sf")



coastline_proj <- st_transform(coastline, 3857)   
points_proj <- st_transform(points_sf, 3857)

```

Calculate distances from coastlines.

```{r}

distances <- st_distance(points_proj, coastline_proj)  # matrix [points x coastline lines]
min_dists <- apply(distances, 1, min) 

```

Put distances as minimum distances.

```{r}

points_sf$dist_km <- as.numeric(min_dists) / 1000
filtered_points <- points_sf %>% filter(dist_km <= 400)


filtered_points <- filtered_points %>%
  select(-timezone) %>%
  select(-OBJECTID)

```

Let's graph 2020 April

```{r}
filtered_2020_4 <- filtered_points %>%
  filter(year == 2020) %>%
  filter(month == 4) 


ggplot() +
  geom_sf(data = coastline, color = "blue") +
  geom_sf(data = filtered_2020_4, aes(color = Anomaly), size = 4) +
  theme_minimal() +
  labs(title = "Points within 400 km of Coastline, April 2020")
```

And let's graph April 2000

```{r}
filtered_2000_4 <- filtered_points %>%
  filter(year == 2000) %>%
  filter(month == 4) 


ggplot() +
  geom_sf(data = coastline, color = "blue") +
  geom_sf(data = filtered_2000_4, aes(color = Anomaly), size = 4) +
  theme_minimal() +
  labs(title = "Points within 400 km of Coastline, April 2000")
```

Change how the dataset looks

```{r}

filtered_points_wide <- filtered_points %>%
  mutate(
    year_month = sprintf("%d-%02d-1", year, month)  # e.g., "1990_01", "1990_02"
  ) %>%
  group_by(geometry, year_month) %>%
  summarize(mean_anomaly = mean(Anomaly, na.rm = TRUE), .groups = "drop") %>%
  pivot_wider(
    names_from = year_month,
    values_from = mean_anomaly,
    names_sort = TRUE
  )


```


Save dataset as RDS

```{r}


saveRDS(filtered_points_wide, file = "/Users/student/Desktop/Data+/climate_plus_summer_2025/EthanDashboardCopy/wide_global_temp_anomaly_1990_2025.rds")

```


Let's graph one data point across time and see the trend

```{r}

coords <- st_coordinates(filtered_points)

data <- filtered_points %>%
  mutate(X = coords[,1], Y = coords[,2]) %>%
  filter(abs(X - -32.5) < 1e-6, abs(Y - 82.5) < 1e-6)

data <- data %>%
  mutate(date = ymd(paste(year, month, 1, sep = "-")))


ggplot(data, aes(x = date, y = Anomaly)) +
  geom_line() +              # line plot over time
  geom_point() +             # points for each month
  labs(title = "Anomaly over Time",
       x = "Date",
       y = "Anomaly") +
  theme_bw()



```

Let's save original filtered points file

```{r}


saveRDS(filtered_points, file = "/Users/student/Desktop/Data+/climate_plus_summer_2025/EthanDashboardCopy/long_global_temp_anomaly_1990_2025.rds")

```

