---
title: "ndGAINdata"
output: html_document
date: "2025-06-23"
---

```{r}
library(shiny)
library(leaflet)
library(dplyr)
library(sf)
library(viridis)
library(tidyverse)
library(qs)
library(pryr)
library(rnaturalearth)
```


```{r setup, include=FALSE}

gain <- read.csv("/Users/student/Desktop/Data+/climate_plus_summer_2025/Data/Primary Datasets/ND GAIN/resources/gain/gain.csv")

gain_long <- gain %>%
  pivot_longer(
    cols = starts_with("X"),
    names_to = "Year",
    values_to = "Value"
  ) %>%
  mutate(Year = as.integer(sub("X", "", Year)))

```

Find for Korea, Australia, Cape Verde, Mexico, and Uganda

```{r}

gain_long_filtered <- gain_long %>%
  filter(ISO3 %in% c("KOR", "AUS", "CPV", "MEX", "UGA", "USA", "VNM"))

```

Graph the findings

```{r}


ggplot(gain_long_filtered, aes(x = Year, y = Value, color = ISO3)) +
  geom_line(size = 1) +
  labs(
    title = "GAIN Over Time for Countries (1995-2022)",
    x = "Year",
    y = "Value",
    color = "Country"
  ) +
  theme_bw()


```

ID_HEAL_01: projected change of deaths from climate change induced diseases

Calculation: The projected change is the percent increase of DALYs from the historical baseline
(2000) to the 2030 estimation using S550 emission scenario.

An indication of the climate change impacts on several types of diseases. The
indicator is a model-based estimate of the quality-adjusted loss of life years under several
different climate scenarios. Disability adjusted life year (DALY) due to malaria, an indication of
the climate change impacts on vector borne diseases, is excluded because more specific
models have been used to project such impacts and it is assessed by another ND-GAIN
indicator, the projected change of length of transmission season of vector-borne diseases (see
below).

```{r}

id_heal_01_input <- read.csv("/Users/student/Desktop/Data+/climate_plus_summer_2025/Data/Primary Datasets/ND GAIN/resources/indicators/id_heal_01/input.csv")
id_heal_01_raw <- read.csv("/Users/student/Desktop/Data+/climate_plus_summer_2025/Data/Primary Datasets/ND GAIN/resources/indicators/id_heal_01/raw.csv")
id_heal_01_raw0 <- read.csv("/Users/student/Desktop/Data+/climate_plus_summer_2025/Data/Primary Datasets/ND GAIN/resources/indicators/id_heal_01/raw0.csv")
id_heal_01_score <- read.csv("/Users/student/Desktop/Data+/climate_plus_summer_2025/Data/Primary Datasets/ND GAIN/resources/indicators/id_heal_01/score.csv")

View(id_heal_01_input)
View(id_heal_01_raw)
View(id_heal_01_raw0)
View(id_heal_01_score)

```




