---
title: "newGPWEDA"
output: html_document
date: "2025-06-06"
---

```{r setup, include=FALSE}

library(dplyr)
library(shiny)
library(leaflet)
library(dplyr)
library(sf)
library(rnaturalearth)
library(bslib)
library(bsicons)
library(ggplot2)
library(plotly)
library(data.table)
library(corrplot)

new_gpw = readRDS("/Users/student/Desktop/Data+/climate_plus_summer_2025/Data/Primary Datasets/gpw_filtered5k.rds")
new_gpw <- new_gpw %>%
  select(COUNTRY, ISOALPHA, everything())

```

Look for missing countries (there are more but here are some):
- Greenland (0) --> MISSING --> 100
- Norway (0) --> MISSING --> 430
- Svalbard (0) --> MISSING --> 1
- Libya (0) --> MISSING --> 22
- Qatar (0) --> MISSING --> 90
- Puerto Rico (0) --> MISSING --> 76112
- Kuwait (0) --> MISSING --> 6
- Tonga (0) --> MISSING --> 5

- U.S.A (794787) --> (794787 / 950831)
- Cuba (93)
- Saudi Arabia (26)
- DPRK (55)
- Egypt (103)
- Eritrea (2)
- Burma (129)
- Angola (22)
- Taiwan (342)
- UAE (5)
- Gabon (9)
- Sudan (6)
- Timor-Leste (211)
- Guyana (57)

```{r}

greenland <- new_gpw %>%
  filter(ISOALPHA == "GRL")
norway <- new_gpw %>%
  filter(ISOALPHA == "NOR")
cuba <- new_gpw %>%
  filter(ISOALPHA == "CUB")
svalbard <- new_gpw %>%
  filter(ISOALPHA == "SJM")
usa <- new_gpw %>%
  filter(ISOALPHA == "USA")
saudi <- new_gpw %>%
  filter(ISOALPHA == "SAU")
dprk <- new_gpw %>%
  filter(ISOALPHA == "PRK")
egypt <- new_gpw %>%
  filter(ISOALPHA == "EGY")
eritrea <- new_gpw %>%
  filter(ISOALPHA == "ERI")
burma <- new_gpw %>%
  filter(ISOALPHA == "MMR")
libya <- new_gpw %>%
  filter(ISOALPHA == "LBY")
angola <- new_gpw %>%
  filter(ISOALPHA == "AGO")
taiwan <- new_gpw %>%
  filter(ISOALPHA == "TWN")
uae <- new_gpw %>%
  filter(ISOALPHA == "ARE")
gabon <- new_gpw %>%
  filter(ISOALPHA == "GAB")
sudan <- new_gpw %>%
  filter(ISOALPHA == "SDN")
timor_leste <- new_gpw %>%
  filter(ISOALPHA == "TLS")
qatar <- new_gpw %>%
  filter(ISOALPHA == "QAT")
puerto <- new_gpw %>%
  filter(ISOALPHA == "PRI")
kuwait <- new_gpw %>%
  filter(ISOALPHA == "KWT")
guyana <- new_gpw %>%
  filter(ISOALPHA == "GUY")
tonga <- new_gpw %>%
  filter(ISOALPHA == "TON")

```

Let's try mapping out some of the ones that have a lot of points.

```{r}
bounding_boxes <- do.call(rbind, lapply(1:nrow(world), function(i) {
  bb <- st_bbox(world[i, ])
  data.frame(
    name = world$admin[i],
    xmin = bb["xmin"],
    ymin = bb["ymin"],
    xmax = bb["xmax"],
    ymax = bb["ymax"]
  )
}))
```


```{r}

ggplot(taiwan) + 
  geom_sf(data = world, fill = "#ADD8E6", color = "white") +
  geom_sf() +
  coord_sf(xlim = c(120.1062, 121.9512), ylim = c(21.97057, 25.29546), expand = FALSE)

```
```{r}
original_taiwan <- read.csv("/Users/student/Downloads/gpw-v4-admin-unit-center-points-population-estimates-rev11_twn_csv (1)/gpw_v4_admin_unit_center_points_population_estimates_rev11_twn.csv")

original_dprk <- read.csv("/Users/student/Downloads/gpw-v4-admin-unit-center-points-population-estimates-rev11_prk_csv (1)/gpw_v4_admin_unit_center_points_population_estimates_rev11_prk.csv")
```


```{r}
library(data.table)
```

```{r}


new_gpw <- original_global_csv

greenland <- new_gpw %>%
  filter(ISOALPHA == "GRL")
norway <- new_gpw %>%
  filter(ISOALPHA == "NOR")
cuba <- new_gpw %>%
  filter(ISOALPHA == "CUB")
svalbard <- new_gpw %>%
  filter(ISOALPHA == "SJM")
usa <- new_gpw %>%
  filter(ISOALPHA == "USA")
saudi <- new_gpw %>%
  filter(ISOALPHA == "SAU")
dprk <- new_gpw %>%
  filter(ISOALPHA == "PRK")
egypt <- new_gpw %>%
  filter(ISOALPHA == "EGY")
eritrea <- new_gpw %>%
  filter(ISOALPHA == "ERI")
burma <- new_gpw %>%
  filter(ISOALPHA == "MMR")
libya <- new_gpw %>%
  filter(ISOALPHA == "LBY")
angola <- new_gpw %>%
  filter(ISOALPHA == "AGO")
taiwan <- new_gpw %>%
  filter(ISOALPHA == "TWN")
uae <- new_gpw %>%
  filter(ISOALPHA == "ARE")
gabon <- new_gpw %>%
  filter(ISOALPHA == "GAB")
sudan <- new_gpw %>%
  filter(ISOALPHA == "SDN")
timor_leste <- new_gpw %>%
  filter(ISOALPHA == "TLS")
qatar <- new_gpw %>%
  filter(ISOALPHA == "QAT")
puerto <- new_gpw %>%
  filter(ISOALPHA == "PRI")
kuwait <- new_gpw %>%
  filter(ISOALPHA == "KWT")
guyana <- new_gpw %>%
  filter(ISOALPHA == "GUY")
tonga <- new_gpw %>%
  filter(ISOALPHA == "TON")


```

Let's map Puerto Rico

-67.24243
17.94655
-65.591
18.5206

```{r}

puerto <- st_as_sf(puerto, 
                   coords = c("CENTROID_X", "CENTROID_Y"), crs = 4326)

ggplot(puerto) + 
  geom_sf(data = world, fill = "#ADD8E6", color = "white") +
  geom_sf() +
  coord_sf(xlim = c(-67.24243, -65.591), ylim = c(17.94655, 18.5206), expand = FALSE)

```

