---
title: "averageGPWinequityDataEDA"
output: html_document
date: "2025-06-03"
---
```{r}
library(dplyr)
library(shiny)
library(leaflet)
library(dplyr)
library(sf)
library(rnaturalearth)
library(bslib)
library(bsicons)
library(ggplot2)
library(plotly)
library(data.table)
library(corrplot)

```

Load average data and the bounding boxes dataset.

```{r}
# NEW DATA EDA

average <- readRDS("/Users/student/Desktop/Data+/climate_plus_summer_2025/Data/gpw_inequity_average.rds") %>%
  st_transform(4326)

bounding_boxes <- do.call(rbind, lapply(1:nrow(world), function(i) {
  bb <- st_bbox(world[i, ])
  data.frame(
    name = world$admin[i],
    xmin = bb["xmin"],
    ymin = bb["ymin"],
    xmax = bb["xmax"],
    ymax = bb["ymax"]
  )
}))
```

Let's look into Japan:

```{r}

jpn <- filter(average, ISOALPHA == "JPN")
ggplot(jpn) +
  ggtitle(label = "Japan Points Distribution") +
  geom_sf(data = world, fill = "#ADD8E6", color = "white") +
  geom_sf() +
  coord_sf(xlim = c(129.4085, 145.5431), ylim = c(31.02958, 45.55148), expand = FALSE) +
  theme_minimal()
```

Let's look into potential correlations between data.

```{r}
jpn_nogeo <- st_drop_geometry(jpn)
selected_vars <- jpn_nogeo[, 100:120]
cor_matrix <- cor(selected_vars, use = "complete.obs")
```

This checks to see which variables are the same across the entire country

```{r}
zero_var_cols <- sapply(selected_vars, function(x) sd(x, na.rm = TRUE) == 0) 
zero_var_cols
```
Looks like the only ones we CAN use in the correlation matrix are:
- distance_to_coast_km
- mean.count.grav.V2.log.sc 
- povmap.grdi.v1.sc
- perc.pop.world.coastal.merit.10m.log.sc
- vulnerab.score.rank
- hierachical.score.rank.ineq 

```{r}
selected_vars <- jpn_nogeo[, c("distance_to_coast_km", 
                               "mean.count.grav.V2.log.sc",
                               "povmap.grdi.v1.sc",
                               "perc.pop.world.coastal.merit.10m.log.sc")]
cor_matrix <- cor(selected_vars, use = "complete.obs")
corrplot(cor_matrix, method = "color", type = "upper", tl.cex = 0.8, 
         title = "Correlation Matrix for Japan")
```
```{r}
print(cor_matrix)
```
Looks like the ones that are actually interesting are:
- povmap.grdi.v1.sc & perc.pop.world.coastal.merit.10m.log.sc (-0.75261687)

Let's try graphing some.

```{r}

variables <- c("vulnerab.score.rank", "hierachical.score.rank.ineq", "povmap.grdi.v1.sc", "perc.pop.world.coastal.merit.10m.log.sc")

```

This accidentally ended up being for the whole entire world
```{r}
plot(df$povmap.grdi.v1.sc, df$perc.pop.world.coastal.merit.10m.log.sc)
```
```{r}
average_nogeo <- st_drop_geometry(average)
selected_vars <- average_nogeo[, c("distance_to_coast_km", 
                               "mean.count.grav.V2.log.sc",
                               "povmap.grdi.v1.sc",
                               "perc.pop.world.coastal.merit.10m.log.sc")]
cor_matrix <- cor(selected_vars, use = "complete.obs")
corrplot(cor_matrix, method = "color", type = "upper", tl.cex = 0.8, 
         title = "Correlation Matrix for World")

```
```{r}
average_nogeo <- st_drop_geometry(average)
selected_vars <- average_nogeo[, c("vulnerab.score.rank", "hierachical.score.rank.ineq", "gov.score.rank", "ineq.score.rank" )]
cor_matrix <- cor(selected_vars, use = "complete.obs")
corrplot(cor_matrix, method = "color", type = "upper", tl.cex = 0.8, 
         title = "Correlation Matrix for World")
```

```{r}
print(cor_matrix)
```



```{r}
for (x_var in variables) {
  for (y_var in variables) {
    plot(df[[x_var]], df[[y_var]],
         xlab = x_var,
         ylab = y_var,
         main = paste(x_var, "vs", y_var))
  }
}
```
This one can actually be just for Japan:

```{r}
for (x_var in variables) {
  for (y_var in variables) {
    plot(jpn[[x_var]], jpn[[y_var]],
         xlab = x_var,
         ylab = y_var,
         main = paste(x_var, "vs", y_var))
  }
}
```

Let's try for Australia.

```{r}
aus <- filter(average, ISOALPHA == "AUS")
ggplot(aus) +
  ggtitle(label = "Australia Points Distribution") +
  geom_sf(data = world, fill = "#ADD8E6", color = "white") +
  geom_sf() +
  coord_sf(xlim = c(113.33895308, 153.569469), ylim = c(-43.6345973, -10.6681857), expand = FALSE) +
  theme_minimal()
```
```{r}
aus_nogeo <- st_drop_geometry(aus)
selected_vars <- aus_nogeo[, c("distance_to_coast_km", 
                               "mean.count.grav.V2.log.sc",
                               "povmap.grdi.v1.sc",
                               "perc.pop.world.coastal.merit.10m.log.sc")]
cor_matrix <- cor(selected_vars, use = "complete.obs")
corrplot(cor_matrix, method = "color", type = "upper", tl.cex = 0.8, 
         title = "Correlation Matrix for Australia")
```
```{r}
print(cor_matrix)

```

The one that stands out the most is povmap.grdi.v1.sc vs. perc.pop.world.coastal.merit.10m.log.sc

```{r}

plot(aus$perc.pop.world.coastal.merit.10m.log.sc, aus$povmap.grdi.v1.sc)

```

Relative Deprivation Index (povmap.grdi.v1.sc): Relative levels of multidimensional deprivation and poverty based on local socio-economic and demographic conditions. 

Climate Vulnerability (perc.pop.world.coastal.merit.10m.log.sc): Percentage of the total population that lives below 10 m.